An Investigation of Confusing Code Patterns in JavaScript

### 2022-12-15 response-letter01.tex

\documentclass{letter}

\usepackage{xcolor}

\usepackage{hyperref}
\signature{Rodrigo Bonif\'{a}cio}
\address{Computer Science Department \\ University of Bras\'{i}lia \\ Brazil}
\begin{document}

\begin{letter}{Journal of Systems and Software Editors-in-Chief}

\opening{Dear Profs. P. Avgeriou and D. Shepherd}

First of all, we would like to thank the anonymous reviewers of
our paper.

{\bf R1 concern about the paper's structure}

We changed the structure of the paper. In the previous version,
we organized the paper in terms of Study Settings, Results, and Discussion.
Each one of those sections present information of the four studies.
We now orgnized the paper in terms of one section per study, and each
one of these sections detail the Study Settings and Results. We
hope that this new structure might have improved the paper understanding.
This new structure also makes easier to change the order of the
studies' presentation. So, if the reviewers agree that we should
present our fourth study first (as R1 suggested), we can easily
apply this change. Thanks for the comment.

{\bf R1 concern about using effect size}

We are computing effect size even when the differences are
not statistically relevant because this might reveal practical
relevance, in particular in border lines scenarios---as is the case
that we show in the paper. If the reviewers do not agree with our
decision, we can remove these calculations from the paper in the
final version. Thanks for the comment.

{\bf R2 concern about the relevance of studying atoms of confusion in JavaScript}

We changed the first section of the paper (Introduction, page 3) to motivate our study better. The
changes are in {\color{blue}blue}. Thanks for the comment.

{\bf R2 concern about the procedures we use to identify atom candidates}

We clarified this point on the current section 4.1 of the paper (page 10). The changes are in
{\color{blue}blue}. Thanks for the comment.

{\bf R3 concern about the lengthy of the MSR study}

Considering the new organization of the paper, we have actually devoted almost 5 pages
of the paper to the MSR study on Prevalence (excluding the discussion section). This is comparable to the
other three studies. Nonetheless, if the reviewers agree that we should put more emphasis
on specific parts of this study, we can enrich this section in the final
version of the paper. Thanks for the comment.

{\bf R3 concerns about the lack of replication package}

We are sharing our replication package at Zenodo. {\color{red}Incluir um link aqui}.
We totally agree that we should have shared it before. We hope that our package might
help understand our work better. We have included examples of the CodeQL queries we used
to search for atoms of confusion in open source JavaScript projects---both in the paper
as well as in the replication package. The replication package also contains all atoms of
confusion we collected using the CodeQL queris (MSR study) and the atoms candidates
we explored in the survey and experiments. Some of our queries might lead to
either false positives and false negatives, although we worked hard to improve
accuracy. We include a new threat to validity regarding this issue, that might threaten
any comparison of atoms prevalence in JavaScript and C/C++. Thanks again for this comment.

{\bf R3 concerns about the presentation of both experiments (RM/LS)}

We justify better the report of both experiments in Section 3 (page 7). The changes
are in {\color{blue}blue}. We revisited this issue again on Section 8.3 (pages 36, 37, and
38). The changes are in {\color{blue}blue}.

Unfortunatly, we cannot use the same test (Chi-square) for both experiments. The McNemar test targets paired,
dependent samples (\url{https://en.wikipedia.org/wiki/McNemar_test}). This is the case of the repeated measures study.
Chi-square is for independent samples. The Chi-square test assumes that the groups are not related
(\url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3900058/}). This is precisely what happens in the Latin square study,
where each participant is exposed to a single treatment. This is not the case in the repeated measures study,
where the same subjects receive both treatments.

We have also fixed all minor issues the reviewers reported. Thanks one more time.

\closing{Yours Faithfully,}

\end{letter}
\end{document}

### 2022-12-16 s02.tex
improving threats to validity

 % latex table generated in R 4.0.4 by xtable 1.8-4 package
 % Thu Apr 22 07:58:21 2021
 \begin{table*}[ht]
\caption{Hypotheses Testing (misunderstanding and time). Asterisks ($^{*}$) indicate a statistically significant difference.}
 \centering
 {\scriptsize
 \begin{tabular}{lrrr|rr}
   \toprule
      & \multicolumn{3}{c}{Correcness analysis} & \multicolumn{2}{c}{Time analysis} \\ \toprule
Atom  & Chi-square & Odds           & Confidence & Mann-Whitey& Cliff's\\ 
 &  test &Ratio          & Interval &  U test & Delta \\  \midrule
Comma Operator & \textbf{$<$ 0.0001*} & 19.02 & (6.60, 68.22) & \textbf{$<$ 0.0001*} & -0.53 \\ 
Automatic Semicolon Insertion & \textbf{$<$ 0.0001*} & 39.33 & (9.21, 356.62) & 0.0980 & -0.16 \\ 
Post-Increment & \textbf{0.0015*} & 4.83 & (1.73, 15.74) & 0.1201 & 0.15 \\ 
Omitted Curly Braces & 0.0510 & 2.35 & (1.00, 5.77) & 0.9253 & -0.01 \\ 
Assignment as Value & \textbf{0.0035*} & 8.39 & (1.81, 79.12) & 0.8268 & 0.02 \\ 
Implicit Predicate & \textbf{0.0112*} & 6.95 & (1.46, 66.56) & \textbf{0.0029*} & -0.29 \\ 
Logic as Control Flow & 0.2920 & 1.54 & (0.73, 3.28) & \textbf{$<$ 0.0001*} & -0.39 \\ 
Conditional Operator & 0.1590 & 2.73 & (0.74, 12.57) & 0.2407 & -0.12 \\ 
Pre-Increment & 0.7015 & 1.25 & (0.55, 2.85) & \textbf{0.0062*} & 0.27 \\ 
Arithmetic as Logic & 1 & 0.84 & (0.22, 3.12) & \textbf{0.0172*} & 0.23 \\ 
    \bottomrule
 \end{tabular}}
 \label{tab:hypothesis-testing}
 \end{table*}

>>>>

 % latex table generated in R 4.0.4 by xtable 1.8-4 package
 % Thu Apr 22 07:58:21 2021
 \begin{table*}[ht]
\caption{Hypotheses Testing (misunderstanding and time). Asterisks ($^{*}$) indicate a statistically significant difference.}
 \centering
 {\scriptsize
 \begin{tabular}{lrrr|rr}
   \toprule
      & \multicolumn{3}{c}{Correctness analysis} & \multicolumn{2}{c}{Time analysis} \\ \toprule
Atom  & Chi-square & Odds           & Confidence & Mann-Whitey& Cliff's\\ 
 &  test &Ratio          & Interval &  U test & Delta \\  \midrule
Comma Operator & \textbf{$<$ 0.0001*} & 19.02 & (6.60, 68.22) & \textbf{$<$ 0.0001*} & -0.53 \\ 
Automatic Semicolon Insertion & \textbf{$<$ 0.0001*} & 39.33 & (9.21, 356.62) & 0.0980 & -0.16 \\ 
Post-Increment & \textbf{0.0015*} & 4.83 & (1.73, 15.74) & 0.1201 & 0.15 \\ 
Omitted Curly Braces & 0.0510 & 2.35 & (1.00, 5.77) & 0.9253 & -0.01 \\ 
Assignment as Value & \textbf{0.0035*} & 8.39 & (1.81, 79.12) & 0.8268 & 0.02 \\ 
Implicit Predicate & \textbf{0.0112*} & 6.95 & (1.46, 66.56) & \textbf{0.0029*} & -0.29 \\ 
Logic as Control Flow & 0.2920 & 1.54 & (0.73, 3.28) & \textbf{$<$ 0.0001*} & -0.39 \\ 
Conditional Operator & 0.1590 & 2.73 & (0.74, 12.57) & 0.2407 & -0.12 \\ 
Pre-Increment & 0.7015 & 1.25 & (0.55, 2.85) & \textbf{0.0062*} & 0.27 \\ 
Arithmetic as Logic & 1 & 0.84 & (0.22, 3.12) & \textbf{0.0172*} & 0.23 \\ 
    \bottomrule
 \end{tabular}}
 \label{tab:hypothesis-testing}
 \end{table*}

### 2022-12-16 threats.tex
changing the paper organization
improving threats to validity

@@

\textbf{External validity.} Our results suggest that the selected idioms and code constructs may lead to confusion for small code snippets, but it is not clear if that result extrapolates to other scenarios. Even though it is likely that in larger code bases the confusion induced by these constructs may be even greater, the existence of additional context may mitigate this effect. Another possible threat to the generalizability of our results, in particular for the survey and interviews, lies in the fact that the analyzed atoms may rarely occur in real JavaScript code bases. To mitigate that threat, we have analyzed 72 popular open source JavaScript repositories and found out that most of them are common, occurring at least once per 1,000 lines of code. Only Arithmetic as Logic and Comma Operator occur less frequently than once per 10,000 lines of code. 
% \castor{A forma como os átomos foram detectados é uma ameaça que não discutimos.}

>>>>

In this section we discuss some of the main threats to the validity of this work.

@@

\textbf{External validity.} Our results suggest that the selected idioms and code constructs may lead to confusion for small code snippets, but it is not clear if that result extrapolates to other scenarios. Even though it is likely that in larger code bases the confusion induced by these constructs may be even greater, the existence of additional context may mitigate this effect. Another possible threat to the generalizability of our results, in particular for the survey and interviews, lies in the fact that the analyzed atoms may rarely occur in real JavaScript code bases. To mitigate that threat, we have analyzed 72 popular open source JavaScript repositories and found out that most of them are common, occurring at least once per 1,000 lines of code. Only Arithmetic as Logic and Comma Operator occur less frequently than once per 10,000 lines of code.

\revised{\textbf{Precision of our CodeQL queries.}
  We relied on existing CodeQL JavaScript predicates to implement most of our source code queries. However,
  we still had to implement our own predicates for three specific queries: the queries that search for the atoms Automatic Semicolon Insertion, Comma Operator, and Omitted Curly Braces.
  Figure~\ref{lst:codeql-automatic} shows examples of predicates we have implemented. Our test cases and manual assessments revealed that our queries correctly identified the
  atoms in the source code. Unfortunately, we observed that a predefined CodeQL predicate (named mayHaveBooleanValue) we used to mine the Implicit Predicate atom might have
  produced many false positives. That is, there are situations where this predicate wrongly states that an expression does not have a boolean value.
  This issue might have inflated the number of Implicit Predicate occurrences we reported in Section~\ref{sec:s04}}.

\begin{figure}[htb]
\begin{lstlisting}[language=CodeQL, basicstyle=\scriptsize]
import javascript

predicate hasTokenAfter(ReturnStmt r) {
  r.getLastToken().getNextToken().getValue() != "}"
}

predicate isNextTokenVariable(ReturnStmt r) {
  r.getLastToken().getNextToken() instanceof IdentifierToken and 
  r.getLastToken().getNextToken().getNextToken().getValue() != "="
}

predicate isNextTokenLiteral(ReturnStmt r) {
  ( r.getLastToken().getNextToken() instanceof StringLiteralToken or
    r.getLastToken().getNextToken() instanceof NumericLiteralToken ) and 
  r.getLastToken().getNextToken().getNextToken().getValue() != "="
 }

from ReturnStmt r
where not r.getTopLevel().isMinified() and
 r.isSubjectToSemicolonInsertion() and 
 r.getNumChildStmt() = 0 and
 r.getNumChildExpr() = 0 and
 not r.toString().matches("%;") and
 hasTokenAfter(r) 
select r, r.getLocation().getStartLine(), r.getLocation().getFile()
\end{lstlisting}
  \caption{CodeQL query that searches for the atom Automatic Semicolon Insertion. Note that, for this query,
    we had to implement three CodeQL predicates.}
  \label{lst:codeql-automatic}
\end{figure}
% \castor{A forma como os átomos foram detectados é uma ameaça que não discutimos.}
